---
title: "Swift Key Coursera NLP Project"
output: html_notebook
---

# Description 

## Swiftkey is a Natural Language processing project on Final Capstone project in the Coursera Data Science Specialization Course

###Tasks to accomplish

  - Tokenization - identifying appropriate tokens such as words, punctuation, and numbers. Writing a function that takes a file as input and returns a tokenized version of it.
  - Profanity filtering - removing profanity and other words you do not want to predict.

```{r Loading the Data, message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
path1 <- "C:/Public/Data-Science[JHU-Coursera]/project-nlp"

for(i in list.files(path = path1, pattern = "\\.zip$")) {
  unzip(i, overwrite = TRUE)
}
```

## Getting the Filenames and also getting connections to read the Text files

```{r checking the files, message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
foldernames <- c("DE", "US", "FI", "RU")
filenames_n<-character(0)
for (j in list.dirs(path = path1)) {
  if (length(unlist(strsplit(basename(j),"_")[[1]])) == 2) {
    for (k in list.files(path = j)) {
      # Grep based on Categories and then read
      filenames_n <- c(filenames_n, unlist(strsplit(k, "\\.txt"))[1])
  }
  }
}
```

## Getting the lines

```{r checking the first few lines, message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}
filenames_n

for (j in list.dirs(path = path1)) {
  if (length(unlist(strsplit(basename(j),"_")[[1]])) == 2) {
    for (k in list.files(path = j)) {
      # Grep based on Categories and then read
      print(k)
      dataset:con <- file(k, "r") 
      readLines(con, 1)
  }
  }
}

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
